Demo arXiv template

### 2022-11-23 paper/paper.qmd
add a plot to distinguish variable transformation and scaling

---
title: "Demo arXiv template"
format:
  arxiv-pdf:
    keep-tex: true  
pdf-engine: pdflatex
author:
  - name: H. Sherry Zhang
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, VIC
        country: Australia
    orcid: 0000-0002-7122-1463
    email: huize.zhang@monash.edu
    url: https://huizezhangsh.netlify.app/
  - name: Collaborators
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, VIC
        country: Australia
abstract: |
  - indexes, useful, quantify severity, early monitoring, 
  - A huge number of indexes have been proposed by domain experts, however, a large majority of them are not being adopted, reused, and compared in research or in practice. 
  - One of the reasons for this is the plenty of indexes are quite complex and there is no obvious easy-to-use implementation to apply them to user's data. 
  - The paper describes a general pipeline framework to construct indexes from spatio-temporal data, 
  - This allows all the indexes to be constructed through a uniform data pipeline and different indexes to vary on the details of each step in the data pipeline and their orders. 
  - The pipeline proposed aim to smooth the workflow of index construction through breaking down the complicated steps proposed by various indexes into small building blocks shared by most of the indexes.
  - The framework will be demonstrated with drought indexes as examples, but appliable in general to environmental indexes constructed from multivariate spatio-temporal data
keywords:
  - indexes
  - data pipeline
  - software design
bibliography: bibliography.bib  
editor_options: 
  chunk_output_type: inline
---

@@

@fig-pipeline-steps presents the index construction pipeline from the raw data to the final indexes. 
The available intermediate steps include temporal processing, spatial aggregation, variable transformation, scaling, dimension reduction, normalising, benchmarking, and simplification. 
These steps do not have to be arranged in sequence and not all the steps are required to present in every index construction. 
Spatial aggregation and temporal processing are operations of the data in space and time 
Variable transformation, scaling, and dimension reduction concerns operations on the multivariate aspect of the data.
Normalisation, benchmarking, and simplification can be viewed as the "post-processing" of indexes, which addresses the cross-comparison across different indexes and time, and index communication.

@@

Sometimes, scaling, as a specific type of variable transformation, is merged into dimension reduction as a pre-step, i.e. transform the covariance matrix into correlation matrix before applying the Principal Component Analysis (PCA), but it worth to be separated into its own step to make it explicit.
Its main different from variable transformation is that it can be written in the form of $$[x(s_i;t_j) - \alpha]/\gamma.$$ 
For example, In a z-score standardisation, $\alpha = \bar{x}(s; t)$ and $\gamma = \sigma(s; t)$ are used to standardise the data by the its mean and standard deviation across all time and all space units
Another example is to scale the variable into a unit interval, which can be written as $\alpha = \min[x(s_i, t_j)]$ and $\gamma = \max[x(s_i, t_j)] - \min[x(s_i, t_j)]$.
While being a common technique in data pre-processing, scaling has an impact on the outliers of the data and analysts need to be cautious on checking its effect.

@@

>>>>

---
title: "Demo arXiv template"
format:
  arxiv-pdf:
    keep-tex: true  
pdf-engine: pdflatex
author:
  - name: H. Sherry Zhang
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, VIC
        country: Australia
    orcid: 0000-0002-7122-1463
    email: huize.zhang@monash.edu
    url: https://huizezhangsh.netlify.app/
  - name: Collaborators
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, VIC
        country: Australia
abstract: |
  - indexes, useful, quantify severity, early monitoring, 
  - A huge number of indexes have been proposed by domain experts, however, a large majority of them are not being adopted, reused, and compared in research or in practice. 
  - One of the reasons for this is the plenty of indexes are quite complex and there is no obvious easy-to-use implementation to apply them to user's data. 
  - The paper describes a general pipeline framework to construct indexes from spatio-temporal data, 
  - This allows all the indexes to be constructed through a uniform data pipeline and different indexes to vary on the details of each step in the data pipeline and their orders. 
  - The pipeline proposed aim to smooth the workflow of index construction through breaking down the complicated steps proposed by various indexes into small building blocks shared by most of the indexes.
  - The framework will be demonstrated with drought indexes as examples, but appliable in general to environmental indexes constructed from multivariate spatio-temporal data
keywords:
  - indexes
  - data pipeline
  - software design
bibliography: bibliography.bib  
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, echo = FALSE)

```

@@

@fig-pipeline-steps presents the index construction pipeline from the raw data to the final indexes. 
The available intermediate steps include temporal processing, spatial aggregation, variable transformation, scaling, dimension reduction, normalising, benchmarking, and simplification. 
These steps do not have to be arranged in sequence and not all the steps are required to present in every index construction. 
Spatial aggregation and temporal processing are operations of the data in space and time. 
Variable transformation, scaling, and dimension reduction concerns operations on the multivariate aspect of the data.
Normalisation, benchmarking, and simplification can be viewed as the post-processing of indexes, which addresses the cross-comparison across different indexes and time, and index communication.

@@

Sometimes, scaling, as a specific type of variable transformation, is merged into dimension reduction as a pre-step, i.e. transform the covariance matrix into correlation matrix before applying the Principal Component Analysis (PCA), but it worth to be separated into its own step to make it explicit.
Its main different from variable transformation is that scaling does not change the shape of the data and it can usually be written in the form of $$[x(s_i;t_j) - \alpha]/\gamma.$$
@fig-scale-var-trans-compare shows a range of operations on a simulated variable. 
For example, In a z-score standardisation, $\alpha = \bar{x}(s; t)$ and $\gamma = \sigma(s; t)$ are used to standardise the data by the its mean and standard deviation across all time and all space units
Another example is to scale the variable into a unit interval, which can be written as $\alpha = \min[x(s_i, t_j)]$ and $\gamma = \max[x(s_i, t_j)] - \min[x(s_i, t_j)]$.
While being a common technique in data pre-processing, scaling has an impact on the outliers of the data and analysts need to be cautious on checking its effect.

@@

```{r fig-scale-var-trans-compare}
#| fig-cap: "Comparison of scaling and variable transformation."
#| fig-height: 8
#| fig-width: 12
library(tidyverse)
library(latex2exp)
set.seed(123)
x <- rgamma(n = 1000, shape = 1, scale = 2)

trans_df <- tibble(origin = x, 
       `z-score` = (x - mean(x)/ sd(x)), 
       qudratic = x^2,
       log = log(x),
       `square root` = sqrt(x),
       `cubic root` = x^(1/3),
       minmax = (x - min(x))/ (max(x) - min(x)),
       boxcox = (x^0.5 - 1)/0.5,
       centering = x - mean(x)
       ) %>% 
  mutate(id = row_number(),
         ) %>% 
  pivot_longer(-id, names_to = "var", values_to = "value") %>% 
  mutate(category = ifelse(
    var %in% c("origin", "z-score", "minmax", "centering"),
    "Scaling", "Var. Trans.")
    )

latex_df <- tibble(
  var = c("origin", "z-score", "qudratic", "log", "square root", "cubic root",
          "minmax", "boxcox", "centering"),
  latex = c(
           r"($x$)",
           r"($\frac{x - \bar{x}}{sd(x)}$)",
           r"($x^2$)",
           r"($log(x)$)",
           r"($\sqrt{x}$)",
           r"($x^{1/3}$)",
           r"($\frac{x - min(x)}{max(x) - min(x)}$)",
           r"($\frac{x^{0.5} - 1}{0.5}$)",
           r"($x - \bar{x}$)"
         )
) %>% 
  left_join(trans_df %>% group_by(var) %>% summarise(max = max(value)))

trans_df %>% 
  ggplot() +
  geom_density(aes(x = value, color = category, y = after_stat(scaled))) + 
  geom_label(data = latex_df, 
             aes(x = 0.8 * max, y = 0.8, label = TeX(latex, output = "character")), 
             parse = TRUE) + 
  facet_wrap(vars(var), scales = "free") + 
  scale_color_brewer(palette = "Dark2") + 
  theme_bw() + 
  theme(legend.position = "bottom")
```

### 2022-11-23 paper/paper.tex
add a plot to distinguish variable transformation and scaling

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, frame hidden, breakable, boxrule=0pt, sharp corners, borderline west={3pt}{0pt}{shadecolor}, enhanced]}{\end{tcolorbox}}\fi

@@

Sometimes, scaling, as a specific type of variable transformation, is
merged into dimension reduction as a pre-step, i.e.~transform the
covariance matrix into correlation matrix before applying the Principal
Component Analysis (PCA), but it worth to be separated into its own step
to make it explicit. Its main different from variable transformation is
that it can be written in the form of \[[x(s_i;t_j) - \alpha]/\gamma.\]
For example, In a z-score standardisation, \(\alpha = \bar{x}(s; t)\)
and \(\gamma = \sigma(s; t)\) are used to standardise the data by the
its mean and standard deviation across all time and all space units
Another example is to scale the variable into a unit interval, which can
be written as \(\alpha = \min[x(s_i, t_j)]\) and
\(\gamma = \max[x(s_i, t_j)] - \min[x(s_i, t_j)]\). While being a common
technique in data pre-processing, scaling has an impact on the outliers
of the data and analysts need to be cautious on checking its effect.

>>>>

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, frame hidden, boxrule=0pt, enhanced, sharp corners]}{\end{tcolorbox}}\fi

@@

Sometimes, scaling, as a specific type of variable transformation, is
merged into dimension reduction as a pre-step, i.e.~transform the
covariance matrix into correlation matrix before applying the Principal
Component Analysis (PCA), but it worth to be separated into its own step
to make it explicit. Its main different from variable transformation is
that scaling does not change the shape of the data and it can usually be
written in the form of \[[x(s_i;t_j) - \alpha]/\gamma.\]
Figure~\ref{fig-scale-var-trans-compare} shows a range of operations on
a simulated variable. For example, In a z-score standardisation,
\(\alpha = \bar{x}(s; t)\) and \(\gamma = \sigma(s; t)\) are used to
standardise the data by the its mean and standard deviation across all
time and all space units Another example is to scale the variable into a
unit interval, which can be written as \(\alpha = \min[x(s_i, t_j)]\)
and \(\gamma = \max[x(s_i, t_j)] - \min[x(s_i, t_j)]\). While being a
common technique in data pre-processing, scaling has an impact on the
outliers of the data and analysts need to be cautious on checking its
effect.

{\centering \includegraphics{paper_files/figure-pdf/fig-scale-var-trans-compare-1.pdf}

\caption{\label{fig-scale-var-trans-compare}Comparison of scaling and
variable transformation.}
